\texttt{SuperScreen}: An open-source package for simulating the magnetic response of two-dimensional superconducting devices

### 2021-11-29 main.tex
overleaf changes

\usepackage{subcaption}
\usepackage[newfloat]{minted}
\usepackage{hyperref}
\usepackage{tcolorbox}
\usepackage{etoolbox}
\BeforeBeginEnvironment{minted}{\begin{center}\begin{tcolorbox}[width=0.9\textwidth]}
\AfterEndEnvironment{minted}{\end{tcolorbox}\end{center}}
% \usemintedstyle[python]{vs}
\usepackage{caption}
\usepackage{xspace}
\usepackage[toc,page]{appendix}

@@

@@

\subsection{Pearl vortices in thin films}
\label{section:examples:pearl-vortices}
Vortices trapped in 2D superconductors ($d\ll\lambda$, where $d$ is the film thickness and $\lambda$ is the London penetration depth), i.e. ``Pearl vortices,'' are associated with different current and magnetic field distributions than Abrikosov vortices trapped in bulk type-II superconductors~\cite{Pearl1964-cl}. The 2D Fourier transform $\tilde{H}_z(\vec{k}, z)$ of the out-of-plane component of the magnetic field $H_z(\vec{r}, z)$ from a Pearl vortex located at the origin $x=y=z=0$ is given by
\begin{equation}
    \tilde{H}_z(\vec{k}, z)=\mathcal{F}\{H_z(\vec{r}, z)\}=\frac{1}{\mu_0}\frac{\Phi_0e^{-|\vec{k}|z}}{1+2\Lambda|\vec{k}|},
    \label{eq:pearl}
\end{equation}
where $\mathcal{F}\{\cdot\}$ is the 2D Fourier transform, $\vec{k}=(k_x, k_y)$ are in-plane spatial frequencies, $z$ is the out-of-plane position at which the field is evaluated, and $2\Lambda = 2\lambda^2 / d$ is the Pearl length~\cite{Pearl1964-cl, Tafuri2004-ap}. The real-space magnetic field distribution near a Pearl vortex can be calculated numerically by taking the inverse Fourier transform of Eq.~\ref{eq:pearl}: $H_z(\vec{r}, z)=\mathcal{F}^{-1}\{\tilde{H}_z(\vec{k}, z)\}$.

To include vortices in a \SuperScreen model, one can simply input a \inline{list} of \inline{superscreen.Vortex} objects when calling \inline{superscreen.solve()}. A \inline{Vortex} object specifies the $x,y$ position for the vortex core, the name of the superconducting layer in which the vortex is pinned, and the number of flux quanta $\Phi_0$ contained in the vortex (which is 1 by default). The field distributions generated by \SuperScreen in the presence of vortices as described in Sections~\ref{section:model:vortices} and~\ref{section:implementation} agree to within a few percent with the distributions obtained using this Fourier transform method, as demonstrated in Figure~\ref{fig:pearl}(c-e). See Ref.~\cite{Brandt2005-wj} for a method to compute the self-energy and interaction energies of vortices in thin films.

@@

Figure~\ref{fig:pearl}(e) shows the fluxoid for a circular region $S$ with radius $r=1\,\um$ enclosing a Pearl vortex trapped in a film as a function of the film's effective penetration depth, $\Lambda$. When $\Lambda=0$, the screening currents decay very quickly away from the center of the vortex, so the ``supercurrent part'' of the $\Phi^f_S$ vanishes. With increasing $\Lambda$, the ``supercurrent part'' accounts for an increasing fraction of the total fluxoid.

While the 2D London model doesn't ``know'' about fluxoid quantization, in the sense that the quantization condition $\Phi^f_S=n\Phi_0$ is not automatically satisfied by solutions to Eq.~\ref{eq:london} for multiply-connected films, we can nevertheless calculate current and field distributions for different fluxoid states in multiply-connected superconductors by adjusting the currents circulating around each hole to realize a prescribed set of fluxoid values. For a structure with $N_h$ holes, we can specify $N_h$ fluxoids $\Phi^f_h$ and find the circulating currents $I_h$ by minimizing the deviation of each fluxoid from its desired value. For $N_h=1$, this is treated as a root-finding problem, which can be solved with typically only three calls to \inline{superscreen.solve()}. When $N_h>1$, it is a least-squares minimization problem with $N_h$ free parameters. This calculation is implemented in the \inline{superscreen.find_fluxoid_solution()} function.

@@

As shown in Ref.~\cite{Brandt2005-wj}, the mutual inductance $M_{ij}$ between holes $i$ and $j$ in a superconducting structure is given by
\begin{equation}
    M_{ij}=\frac{\Phi^f_{S_i}}{I_j},
\end{equation}
where $\Phi^f_{S_i}$ is the fluxoid for a region $S_i$ containing the hole $i$, and $I_j$ is the current circulating around hole $j$. The mutual inductance values for a set of holes form a mutual inductance matrix. The diagonals of the mutual inductance matrix are the hole self-inductances ($M_{ii}=L_i$, the self-inductance of hole $i$), and the matrix is symmetric ($M_{ij}=M_{ji}$) due to the reciprocity theorem. If the penetration depth of the film containing hole $i$ is $\Lambda = 0$, then no field penetrates the film and the fluxoid $\Phi^f_{S_i}$ is equal to $\Phi_i$, the flux through hole $i$. For a device with $N_h$ holes, the $N_h\times N_h$ mutual inductance matrix $\mathbf{M}$ can be computed using the $\inline{Device.mutual_inductance_matrix()}$ method.

@@

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{examples/images/squid-mutuals.pdf}
    \caption{Calculating SQUID susceptometer mutual inductance, defined as the fluxoid induced in the SQUID pickup loop per unit current flowing in the field coil: $M_\mathrm{PL-FC}=\Phi^f_\mathrm{PL} / I_\mathrm{FC}$. Top row (a): Schematics of \SuperScreen models for the field coil and pickup loop region of four sizes of scanning SQUID susceptometer (generated using \inline{Device.draw()}) with pickup loop inner radii ranging from $0.1\,\um$ (``Small'') to $3\,\um$ (``X-Large''). The wiring layers BE, W1, and W2 are shown in blue, orange, and green respectively. Middle row (b): Simulated magnetic field $\mu_0H_z$ evaluated at the plane of layer W1, which contains the pickup loop, normalized by the current $I_\mathrm{FC}$ flowing in the field coil (which is located in layer BE). Bottom row (c): Convergence of the four models, defined as the fractional change in mutual inductance between subsequent iterations $i$. For all four models, the simulated mutual inductance, $M_\mathrm{sim}$, falls within the range of mutual inductance values measured in real devices, $M_\mathrm{exp}$, which were reported in Table 1 of Ref.~\cite{Kirtley2016-zz}. The meshes for the ``Small,'' ``Medium,'' and ``Large'' models consisted of approximately 7,000 vertices and 14,000 triangles, whereas the mesh for the ``X-Large'' model consisted of about 8,000 vertices and 15,000 triangles. The two smaller models converge more quickly than the two larger models for reasons discussed in \ref{appendix:numerics}. Note that for the ``X-Large'' model, we set the thicknesses of layers I1 and I2 both to 400 nm (instead of the nominal values of 150 nm and 130 nm respectively), to ensure convergence (see \ref{appendix:numerics} and Figure~\ref{fig:xlarge-dz-dr}).}
    \label{fig:squid-mutuals}
\end{figure*}

@@

Figure~\ref{fig:squid-mutuals}(a) shows \SuperScreen \inline{Device} models of the field coil and pickup loop region of four types of real SQUID susceptometers, with geometry taken from the layout artwork (GDS) files. In addition to the field coil and pickup loop, there are superconducting shields that limit the amount of magnetic flux that can penetrate the leads connecting the loops to the rest of the circuit (the rest of the circuit is not modeled). There are three relevant layers of superconducting films: the base electrode (BE), which is furthest from the sample contains the field coil; the first wiring layer (W1), which contains the pickup loop and a shield covering the field coil leads; and the second wiring layer (W2), which is closest to the sample and contains a shield covering the pickup loop leads. There are two insulating layers: I1, which separates BE and W1, and I2, which separates W1 and W2.
% Table~\ref{table:squid-layers} shows the thickness $d$ of each layer in the real devices, as well as the vertical position of the 2D films used to model each layer. 
For the superconducting layers, which are made of Nb, we take the London penetration depth to be $\lambda = 80\,\nm$, and the layer thicknesses correspond to the real device design (see Table 1 of Ref.~\cite{Kirtley2016-gt} and Figure 8 of Ref.~\cite{Kirtley2016-zz}). It is important to note that because $\lambda < d$ for all three layers, the susceptometers are not in the 2D limit in which the model described in Section~\ref{section:model} is technically valid. Nevertheless, as shown in Figure~\ref{fig:squid-mutuals}, the field coil - pickup loop mutual inductances, $M_\mathrm{sim}$, computed by \SuperScreen lie within the range of mutual inductances measured in real devices, $M_\mathrm{exp}$ (taken from Ref.~\cite{Kirtley2016-zz} Table 1), for all four sizes of susceptometer, indicating that variation in current density along the thickness of each film is not critical in determining the mutual inductance in this case.\footnote{The mutual inductance values are shown in units of $\Phi_0 / \mathrm{A}$, where $1\,\Phi_0 / \mathrm{A}\approx2.068\times10^{-3}\,\mathrm{pH}$.} The convergence of $M_\mathrm{PL-FC}$ as a function of solver iteration (see Section~\ref{section:model:multilayer}) is shown in Figure~\ref{fig:squid-mutuals}(c).

>>>>

\usepackage{subcaption}
\usepackage[newfloat]{minted}
\usepackage{hyperref}
\usepackage{tcolorbox}
\usepackage{etoolbox}
\BeforeBeginEnvironment{minted}{\begin{center}\begin{tcolorbox}[width=\linewidth]}
\AfterEndEnvironment{minted}{\end{tcolorbox}\end{center}}
% \usemintedstyle[python]{vs}
\usepackage{caption}
\usepackage{xspace}
\usepackage[toc,page]{appendix}

@@

\newenvironment{code-onecol}{\captionsetup{type=listing}}{\hfill}
\SetupFloatingEnvironment{listing}{name=Code Block}
\BeforeBeginEnvironment{code-onecol}{\vspace{10pt}\begin{figure}[t!]}
\AfterEndEnvironment{code-onecol}{\end{figure}}

@@

While the 2D London model doesn't ``know'' about fluxoid quantization, in the sense that the quantization condition $\Phi^f_S=n\Phi_0$ is not automatically satisfied by solutions to Eq.~\ref{eq:london} for multiply-connected films, we can nevertheless calculate current and field distributions for different fluxoid states in multiply-connected superconductors by adjusting the currents circulating around each hole to realize a prescribed set of fluxoid values. For a structure with $N_h$ holes, we can specify $N_h$ fluxoids $\Phi^f_h$ and find the circulating currents $I_h$ by minimizing the deviation of each fluxoid from its desired value. For $N_h=1$, this is treated as a root-finding problem, which can be solved with typically only three calls to \inline{superscreen.solve()}. When $N_h>1$, it is a least-squares minimization problem with $N_h$ free parameters. This calculation is implemented in the \inline{superscreen.find_fluxoid_solution()} function. Code Block~\ref{code:fluxoid} demonstrates how to model a device with one or more holes, each in the $n=0$ fluxoid state, subject to a uniform applied field, and Figure~\ref{fig:fluxoid} shows the field and current distributions for a rectangular superconducting film with $\Lambda=0.25\,\um$, which has one rectangular and one elliptical hole. The least-squares minimization for the model shown in Figure~\ref{fig:fluxoid}, with $N_h=2$, required 18 total calls to \inline{superscreen.solve()}.

\begin{code-onecol}
\begin{minted}[fontsize=\small]{python}
import superscreen as sc
# Assume that we have already created a
# superscreen.Device with one or more
# holes, and generated the mesh.
# Specify the desired fluxoid for each
# hole in the device:
fluxoids = {
    hole_name: 0
    for hole_name in device.holes
}
applied_field = sc.sources.ConstantField(1)
field_units = "mT"

result = sc.find_fluxoid_solution(
    device=device,
    fluxoids=fluxoids,
    applied_field=applied_field,
    field_units=field_units,
)
solution, opt_result = result
\end{minted}
\captionof{listing}{Calculating the field and current distributions for fluxoid states in a multiply-connected superconducting film. Given a \inline{Device} with $N_h\geq 1$ holes and a desired fluxoid $\Phi^f_h$ for each hole, \inline{superscreen.find_fluxoid_solution()} optimizes the current $I_h$ circulating around each hole to realize the desired fluxoid state. The function returns a \inline{tuple} of length 2, the first element being the final optimized \inline{superscreen.Solution} and the second element being an instance of either \inline{scipy.optimize.RootResults} (if $N_h=1$) or \inline{scipy.optimize.OptimizeResult} (if $N_h>1$), which contains information about the optimization that was performed. See Figure~\ref{fig:fluxoid} for an example of the results for a film with two holes, both in the $n=0$ fluxoid state.}
\label{code:fluxoid}
\end{code-onecol}

\begin{figure}[t]
    \centering
    \includegraphics[width=\linewidth]{examples/images/fluxoid.pdf}
    \caption{(a) Magnetic field and (b) current density distributions generated by Code Block~\ref{code:fluxoid} for a rectangular superconducting film with $\Lambda=0.25\,\um$, which has one rectangular and one elliptical hole. The film is subject to a uniform applied out-of-plane field of 1 mT, and both holes are set to be in the $n=0$ fluxoid (i.e. Meissner) state. The resulting ciruclating currents are $I_\text{rectangle} = -1.071$ mA and $I_\text{ellipse} = -1.589$ mA, and the residual fluxoid for each hole is smaller than $10^{-7}\,\Phi_0$. These results were computed using a mesh with approximately 5,000 vertices and 10,000 triangles.}
    \label{fig:fluxoid}
\end{figure}

\subsection{Pearl vortices in thin films}
\label{section:examples:pearl-vortices}
Vortices trapped in 2D superconductors ($d\ll\lambda$, where $d$ is the film thickness and $\lambda$ is the London penetration depth), i.e. ``Pearl vortices,'' are associated with different current and magnetic field distributions than Abrikosov vortices trapped in bulk type-II superconductors~\cite{Pearl1964-cl}. The 2D Fourier transform $\tilde{H}_z(\vec{k}, z)$ of the out-of-plane component of the magnetic field $H_z(\vec{r}, z)$ from a Pearl vortex located at the origin $x=y=z=0$ is given by
\begin{equation}
    \tilde{H}_z(\vec{k}, z)=\mathcal{F}\{H_z(\vec{r}, z)\}=\frac{1}{\mu_0}\frac{\Phi_0e^{-|\vec{k}|z}}{1+2\Lambda|\vec{k}|},
    \label{eq:pearl}
\end{equation}
where $\mathcal{F}\{\cdot\}$ is the 2D Fourier transform, $\vec{k}=(k_x, k_y)$ are in-plane spatial frequencies, $z$ is the out-of-plane position at which the field is evaluated, and $2\Lambda = 2\lambda^2 / d$ is the Pearl length~\cite{Pearl1964-cl, Tafuri2004-ap}. The real-space magnetic field distribution near a Pearl vortex can be calculated by taking the inverse Fourier transform of Eq.~\ref{eq:pearl}: $H_z(\vec{r}, z)=\mathcal{F}^{-1}\{\tilde{H}_z(\vec{k}, z)\}$.

To include vortices in a \SuperScreen model, one can simply input a \inline{list} of \inline{superscreen.Vortex} objects when calling \inline{superscreen.solve()}. A \inline{Vortex} object specifies the $x,y$ position for the vortex core, the name of the superconducting layer in which the vortex is pinned, and the number of flux quanta $\Phi_0$ contained in the vortex (which is 1 by default). The field distributions generated by \SuperScreen in the presence of vortices as described in Sections~\ref{section:model:vortices} and~\ref{section:implementation} agree to within a few percent with the distributions obtained using this Fourier transform method, as demonstrated in Figure~\ref{fig:pearl}(c-e). Figure~\ref{fig:pearl}(e) shows the fluxoid for a circular region $S$ with radius $r=1\,\um$ enclosing a Pearl vortex trapped in a film as a function of the film's effective penetration depth, $\Lambda$. When $\Lambda=0$, the screening currents decay very quickly away from the center of the vortex, so the ``supercurrent part'' of the $\Phi^f_S$ vanishes. With increasing $\Lambda$, the ``supercurrent part'' accounts for an increasing fraction of the total fluxoid. See Ref.~\cite{Brandt2005-wj} for a method to compute the self-energy and interaction energies of vortices in thin films.

@@

@@

As shown in Ref.~\cite{Brandt2005-wj}, the mutual inductance $M_{ij}$ between holes $i$ and $j$ in a superconducting structure is given by
\begin{equation}
    M_{ij}=\frac{\Phi^f_{S_i}}{I_j},
\end{equation}
where $\Phi^f_{S_i}$ is the fluxoid for a region $S_i$ containing the hole $i$, and $I_j$ is the current circulating around hole $j$. The mutual inductance values for a set of holes form a mutual inductance matrix. The diagonals of the mutual inductance matrix are the hole self-inductances ($M_{ii}=L_i$, the self-inductance of hole $i$), and the matrix is symmetric ($M_{ij}=M_{ji}$) due to the reciprocity theorem. If the penetration depth of the film containing hole $i$ is $\Lambda = 0$, then no field penetrates the film and the fluxoid $\Phi^f_{S_i}$ is equal to $\Phi_i$, the flux through hole $i$. For a device with $N_h$ holes, the $N_h\times N_h$ mutual inductance matrix $\mathbf{M}$ can be computed using the $\inline{Device.mutual_inductance_matrix()}$ method. For example, the mutual inductance matrix for the device shown in Figure~\ref{fig:fluxoid} is:
\begin{equation*}
    \mathbf{M} = 
    \begin{pmatrix}
    	10.319 & -1.536 \\
        -1.527 & 7.130
    \end{pmatrix}
    \,\mathrm{pH},
\end{equation*}
where the matrix is indexed as $$\begin{pmatrix}\text{ellipse}\\\text{rectangle}\end{pmatrix},$$
and the ``fractional asymmetry'' of $\mathbf{M}$ is $|M_{01} - M_{10}| / \min(|M_{01}|, |M_{10}|)\approx 0.6\%$.

@@

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{examples/images/squid-mutuals.pdf}
    \caption{Calculating SQUID susceptometer mutual inductance, defined as the fluxoid induced in the SQUID pickup loop per unit current flowing in the field coil: $M_\mathrm{PL-FC}=\Phi^f_\mathrm{PL} / I_\mathrm{FC}$. Top row (a): Schematics of \SuperScreen models for the field coil and pickup loop region of four sizes of scanning SQUID susceptometer (generated using \inline{Device.draw()}) with pickup loop inner radii ranging from $0.1\,\um$ (``Small'') to $3\,\um$ (``X-Large''). The wiring layers BE, W1, and W2 are shown in blue, orange, and green respectively. Middle row (b): Simulated magnetic field $\mu_0H_z$ evaluated at the plane of layer W1, which contains the pickup loop, normalized by the current $I_\mathrm{FC}$ flowing in the field coil (which is located in layer BE). Bottom row (c): Convergence of the four models, defined as the fractional change in mutual inductance between subsequent iterations $i$. For all four models, the simulated mutual inductance, $M_\mathrm{sim}$, falls within the range of mutual inductance values measured in real devices, $M_\mathrm{exp}$, which were reported in Table 1 of Ref.~\cite{Kirtley2016-zz}. The mutual inductance values are shown in units of $\Phi_0 / \mathrm{A}$, where $1\,\Phi_0 / \mathrm{A}\approx2.068\times10^{-3}\,\mathrm{pH}$. The meshes for the ``Small,'' ``Medium,'' and ``Large'' models consisted of approximately 7,000 vertices and 14,000 triangles, whereas the mesh for the ``X-Large'' model consisted of about 8,000 vertices and 15,000 triangles. The two smaller models converge more quickly than the two larger models for reasons discussed in \ref{appendix:numerics}. Note that for the ``X-Large'' model, we set the thicknesses of layers I1 and I2 both to 400 nm (instead of the nominal values of 150 nm and 130 nm respectively), to ensure convergence (see \ref{appendix:numerics} and Figure~\ref{fig:xlarge-dz-dr}).}
    \label{fig:squid-mutuals}
\end{figure*}

@@

Figure~\ref{fig:squid-mutuals}(a) shows \SuperScreen \inline{Device} models of the field coil and pickup loop region of four types of real SQUID susceptometers, with geometry taken from the layout artwork (GDS) files. In addition to the field coil and pickup loop, there are superconducting shields that limit the amount of magnetic flux that can penetrate the leads connecting the loops to the rest of the circuit (the rest of the circuit is not modeled). There are three relevant layers of superconducting films: the base electrode (BE), which is furthest from the sample contains the field coil; the first wiring layer (W1), which contains the pickup loop and a shield covering the field coil leads; and the second wiring layer (W2), which is closest to the sample and contains a shield covering the pickup loop leads. There are two insulating layers: I1, which separates BE and W1, and I2, which separates W1 and W2.
% Table~\ref{table:squid-layers} shows the thickness $d$ of each layer in the real devices, as well as the vertical position of the 2D films used to model each layer. 
For the superconducting layers, which are made of Nb, we take the London penetration depth to be $\lambda = 80\,\nm$, and the layer thicknesses correspond to the real device design (see Table 1 of Ref.~\cite{Kirtley2016-gt} and Figure 8 of Ref.~\cite{Kirtley2016-zz}). It is important to note that because $\lambda < d$ for all three layers, the susceptometers are not in the 2D limit in which the model described in Section~\ref{section:model} is technically valid. Nevertheless, as shown in Figure~\ref{fig:squid-mutuals}, the field coil - pickup loop mutual inductances, $M_\mathrm{sim}$, computed by \SuperScreen lie within the range of mutual inductances measured in real devices, $M_\mathrm{exp}$ (taken from Ref.~\cite{Kirtley2016-zz} Table 1), for all four sizes of susceptometer, indicating that variation in current density along the thickness of each film is not critical in determining the mutual inductance in this case. The convergence of $M_\mathrm{PL-FC}$ as a function of solver iteration (see Section~\ref{section:model:multilayer}) is shown in Figure~\ref{fig:squid-mutuals}(c).
